Meeting the Challenges of Modeling Astrophysical Thermonuclear Explosions: \castro, \maestro, and the \amrex\ Astrophysics Suite

### 2017-10-05 astronum_2017/astronum-amrex.tex
new fig

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{titan_edison_maestro_scaling}
\includegraphics[width=0.48\linewidth]{summitdev_scaling}
\caption{\label{fig:scaling} (left) \maestro\ strong scaling on NERSC edison
and OCLF titan for a 3-d XRB problem.  Two different problem sizes are show.
(right) \starlord\ scaling across nodes on the OLCF summitdev platform, using
up to 4 GPUs per node.}
\end{figure}

@@

Castro scaling...

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{CUDA-Fortran-VODE-Integrator}
\begin{minipage}[b]{0.48\linewidth}
\caption{\label{fig:scaling} (left) \maestro\ strong scaling on NERSC edison
and OCLF titan for a 3-d XRB problem.  Two different problem sizes are show.
(right) \starlord\ scaling across nodes on the OLCF summitdev platform, using
up to 4 GPUs per node.\vspace{2em}}
\end{minipage}
\end{figure}

@@

\section{Science results}

@@

Figure~\ref{fig:current-runs} shows some of our most recent science
simulations.  The left panel is an image of convection in the helium
layer on a sub-Chandra white dwarf.  This is part of a study of the
early stages of the double detonation SNe Ia model.  Using \maestro,
we are able to model the convection in the He layer for many turnover
times and saw a range of outcomes depending on the mass of the white
dwarf and He layer, including a global He-nova-like behavior where the
entire layer runsway together and localized
burning~\cite{subchandra2}.

The middle figure shows convection in a H/He layer on a neutron star,
as a model of the early burning in an XRB.  This \maestro\ model was
the first 3D model of convection for this problem~\cite{xrb3}.

The rightmost image is the coalesced remains of the merger of a
0.9~$M_\odot$ and 0.6~$M_\odot$ WD.

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{titan_edison_maestro_scaling}
\includegraphics[width=0.48\linewidth]{summitdev_scaling} \\
\includegraphics[width=0.48\linewidth]{castro_scaling}
\caption{\label{fig:scaling} (left) \maestro\ strong scaling on NERSC edison
and OCLF titan for a 3-d XRB problem.  Two different problem sizes are show.
(right) \starlord\ scaling across nodes on the OLCF summitdev platform, using
up to 4 GPUs per node.}
\end{figure}

@@

Castro scaling is a placeholder...

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{CUDA-Fortran-VODE-Integrator}
\begin{minipage}[b]{0.48\linewidth}
\caption{\label{fig:cudaode} GPU speed up over single CPU core for two
  reaction networks.  The 13-isotope alpha-chain is a
  potential network for the sub-Chandra simulations we propose here.
  The 9-isotope Urca network is used for our Urca simulations.  A
  variety of grid sizes were used, $16^3$, $32^3$, $64^3$, and
  $128^3$.  In all cases, we see significant speed up on the GPU
  vs.\ a single core.  These speed-up numbers are from our testbed
  workstation with a Titan X Pascal board and Intel Xeon X5650 processors.}
\end{minipage}
\end{figure}

@@

\section{Some science results}

@@

Figure~\ref{fig:current-runs} shows some of our recent science
simulations.  The left panel is an image of convection in the helium
layer on a sub-Chandra white dwarf.  This is part of a study of the
early stages of the double detonation SNe Ia model.  Using \maestro,
we are able to model the convection in the He layer for many turnover
times and saw a range of outcomes depending on the mass of the white
dwarf and He layer, including a global He-nova-like behavior where the
entire layer runsway together and localized
burning~\cite{subchandra2}.

The middle figure shows convection in a H/He layer on a neutron star,
as a model of the early burning in an XRB.  This \maestro\ model was
the first 3D model of convection for this problem~\cite{xrb3}.  This
study showed that the convective field became fully turbulent,
achieving a Kolmogorov spectrum, and the overall dynamics was very
different than our earlier 2-d simulations.

The rightmost image is the coalesced remains of the merger of a
0.9~$M_\odot$ and 0.6~$M_\odot$ WD performed with \castro.  This used
the developments from \cite{wdmergerI}.  Our primary focus with this
suite of simulations is understanding the numerical sensitivity of
mergers and collisions on the burning that takes place.

\section{Summary and future development}

We described our suite of astrophysics codes built on the
\amrex\ framework.  These codes were developed to model problems in
stellar astrophysics spanning from low speed convection to explosive
burning.  A major theme of the codes is the open development model,
with the code development done on github and all problem files needed
to recreate any science results freely available.  Future development
efforts for \maestro\ include higher-order hydrodynamics and
time-integration and rotation.  For \castro, we are investigating
strong coupling between hydrodynamics and reactions, new solvers
(including MHD), and finishing the GPU port.
